### 2016-06-01 Chapters/S1_Table.tex

\documentclass[a4paper,12pt]{article}

\input{suppinfo_preambule}
% Add an S1
\renewcommand{\thefigure}{S2}%

\usepackage{color} 
\newcommand{\tr}[1]{{\color{red} #1}}
\newcommand{\tb}[1]{{\color{blue} #1}}

% Add an S1
\renewcommand{\thetable}{S\arabic{table}}%

\begin{document}
\begin{flushleft}
{\Large
\textbf\newline{\textbf{S1 Table -- Parameter values of self-replicator model\footnote{Supporting Information of "\input{title}"}}}
}
\newline
\input{authors_affil}
\end{flushleft}

\begin{table}[!h]
\centering
\begin{tabular}{|c|c|c|c|}
\hline
Parameter & Unit & Literature value & Fitted value \\
\hline
$\gamma$ & $-$ & No value & 1.39\\
\hline
$k_R$ & h$^{-1}$ & 3.6 & 2.23\\
\hline
$e_M$ for M63+glycerol & h$^{-1}$ &  $< 3.6$ & 0.587\\
\hline
$e_M$ for M63+glucose & h$^{-1}$ &  $< 3.6$ & 0.867\\
\hline
$e_M$ for cAA+glycerol & h$^{-1}$ &  $< 3.6$ & 1.07\\
\hline
$e_M$ for cAA+glucose & h$^{-1}$ &  $< 3.6$ & 1.57\\
\hline
$e_M$ for RDM+glycerol & h$^{-1}$ &  $< 3.6$ & 3.48\\
\hline
$e_M$ for RDM+glucose & h$^{-1}$ &  $< 3.6$ & 4.76\\
\hline
$\beta K_R$ & $-$  & 0.003 & Not fitted\\
\hline
\end{tabular}
\caption{\textbf{Parameter values of self-replicator model}
The parameter values in the model were obtained by fitting Eq.~\ref{eq:alpha_mu_optimal_dim} to the data of Scott et al~\cite{scott_interdependence_2010} (Fig.~\ref{fig:model_validation} in the main text), as described in the \textit{Methods} section. They are compared with order-of-magnitude estimates from the literature (\ref{S2_Text}). }
\end{table}

\bibliographystyle{plos2015}
\bibliography{references}

\end{document}