### 2016-06-01 Chapters/main_plos.tex

\vspace*{0.35in}
\input{titlepage}

% Please keep the abstract below 300 words
\input{abstract}

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step. 
% Author Summary not valid for PLOS ONE submissions.   
\input{author_summary}

\linenumbers

%%%%%% CORE OF THE ARTICLE %%%%%
\input{core}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Methods}

\subsection*{Steady-state analysis of model}
\label{sec:methods_model_analysis}

The nondimensional version of the model, given by Eqs~\ref{eq:pdefnondim}-\ref{eq:rdefnondim}, was used for a steady-state analysis of the self-replicator.
Eqs~\ref{eq:pdefnondim}-\ref{eq:rdefnondim} were derived from the original model of Eqs~\ref{eq:pdef}-\ref{eq:rdef} by means of the following rescalings:
\begin{equation*}
\hat{p}  = \beta \, p,\;\;\;
\hat{r}  = \beta \, r,\;\;\;
\hat{t}  = k_R \, t,\;\;\;
E_M = e_M/k_R,\;\;\;
K = \beta K_R.
\end{equation*}

As shown in \ref{S1_Text}, for a constant environment $E_M$ and constant resource allocation $\alpha$, the system has two steady states: a trivial unstable steady state $(\hat{p}^*, \hat{r}^*) = (0,1)$, allowing no growth in the absence of precursors, and a steady state with a positive growth rate given by
\begin{equation}
\label{eq:meth_steadystate}
(\hat{p}^*, \hat{r}^*) = \left( \frac{(1-\alpha)\, E_M - \alpha + \sqrt{[(1-\alpha)\, E_M - \alpha]^2 + 4\alpha\, (1-\alpha)\, E_M\, K}}{2\alpha}, \alpha \right).
\end{equation}
The two eigenvalues of the Jacobian matrix evaluated at $(\hat{p}^*, \hat{r}^*)$ are negative (\ref{S1_Text}), so that this steady state is stable.

The growth rate at steady state, as a function of $\hat{p}^*$ and $\hat{r}^*$, is given by Eq.~\ref{eq:munondim}, which we repeat here for clarity:
\begin{equation*}
\hat{\mu}^* = \frac{\hat{p}^*}{K + \hat{p}^*} \, \hat{r}^* .
\end{equation*}
Evaluating $d\hat{p}/dt=0$ at $(\hat{p}^*, \hat{r}^*)$ allows $\hat{r}^*$, and therefore $\hat{\mu}^*$, to be written as a function of $\hat{p}^*$ (\ref{S1_Text}).
Accordingly, we can compute $\partial\hat{\mu}^*/\partial\hat{p}^*$ and, when setting this partial derivative to 0, determine the maximum growth rate at steady state $\mu_{opt}^*$ and the optimal resource allocation $\alpha_{opt}^*$ bringing about this maximal growth rate.
As shown in \ref{S1_Text}, $\mu_{opt}^*$ and $\alpha_{opt}^*$ can be written as explicit functions of either the environment $E_M$:
\begin{equation}
\label{eq:alpha_mu_optimal}
\alpha_{opt}^* = \frac{E_M + \sqrt{K\, E_M}}{E_M + 2\sqrt{K\, E_M} + 1}
, \;\;\;\;\;\;\;\;\;\;\;\;\;\; 
\hat{\mu}^*_{opt} = \frac{E_M}{E_M + 2\sqrt{K\, E_M} + 1},
\end{equation}
or the precursor abundance $\hat{p}_{opt}^*$:
\begin{equation}
\label{eq:alpha_mu_optimal_p}
\alpha_{opt}^* = \frac{\hat{p}^{*}_{opt}}{\hat{p}^{*}_{opt} + \frac{K}{K+\hat{p}^{*}_{opt}}(1+\hat{p}^{*}_{opt})}
, \;\;\;\;\;\;\;\;\;\;\;\;\;\; 
\hat{\mu}^*_{opt} = \frac{\hat{p}^{* 2}_{opt}}{\hat{p}^{* 2}_{opt} + 2K\hat{p}^*_{opt} + K}.
\end{equation}
The above equations were used for the derivation of the control strategies (see below).

\subsection*{Model fitting}
\label{sec:methods_model_fitting}

As can be seen by comparing Figs~\ref{fig:model_validation}\textit{A} and~\ref{fig:model_validation}\textit{B}, growth-rate maximization in the self-replicator model leads to a good qualitative correspondence with the growth laws.
In order to determine if a good quantitative fit of the model with the data from Scott~\textit{et al.} \cite{scott_interdependence_2010} can be obtained, for reasonable parameter values, we estimated $e_M$ and $k_R$ in Eqs~\ref{eq:pdef}-\ref{eq:rdef} from the measured RNA/protein mass ratios.
At steady state, the RNA/protein mass ratio can be interpreted as proportional to $\hat{r}^*$ (and thus $\alpha_{opt}^*$), with an unknown (dimensionless) proportionality constant $\gamma$ (see~\cite{scott_interdependence_2010} for details on the use of the RNA/protein mass ratio as a proxy for the ribosomal protein mass fraction):
\begin{equation}
\hat{r}^* = \alpha_{opt}^* = \gamma \, \frac{\text{RNA mass}}{\text{protein mass}}.
\label{eq:gamma}
\end{equation}
Reformulating Eq.~\ref{eq:alpha_mu_optimal} in terms of the original parameters $e_M$ and $k_R$, which have physical dimensions facilitating the biological interpretation of their values, we obtain a straighforward relation between $e_M$, $k_R$, $K$, $\alpha_{opt}^*$ and $\mu_{opt}^*$:
\begin{equation}
\label{eq:alpha_mu_optimal_dim}
\alpha_{opt}^* = \frac{e_M + \sqrt{K\, e_M\, k_R}}{e_M + 2\sqrt{K\, e_M\, k_R} + k_R}
, \;\;\;\;\;\;\;\;\;\;\;\;\;\; 
\mu^*_{opt} = \frac{e_M \, k_R}{e_M + 2\sqrt{K\, e_M\, k_R} + k_R}.
\end{equation}

Eqs~\ref{eq:gamma}-\ref{eq:alpha_mu_optimal_dim} were used to estimate values of $k_R$ and $\gamma$, as well as $e_M$ for each of the six growth conditions, from the measurements of the growth rate and the RNA/protein mass ratio. 
The value $K$ was not estimated from the experimental data, but set to a value inferred from the literature (\ref{S1_Text}).
The optimization process was carried out by means of the differential evolution algorithm of Storn and Price~\cite{storn_differential_1997}.
The results are shown in Fig.~\ref{fig:model_validation}\textit{B}, while the estimated parameter values are summarized in \ref{S1_Table}.
The parameter values are in very good agreement with order-of-magnitude values determined from the literature (\ref{S2_Text} and \ref{S1_Table}).

\subsection*{Solution of optimal control problem}

The optimal control problem of Eq.~\ref{eq:J} consists in identifying the function $\alpha_{opt}(t)$ that maximizes the integral of the growth rate $\hat{\mu}$ over an interval $[0, \tau]$.
In order to solve this problem, we first redefined it over an infinite horizon (\textit{i.e.}, $\tau \rightarrow \infty $) in order to avoid boundary effects occurring over finite time intervals, in particular the depletion of precursors just before reaching $\tau$.
With $\mathcal{U}=\{\alpha:\mathbb{R}^+ \rightarrow [0,1] \}$ the set of admissible controls, the full optimization problem for the nondimensionalized system is given by
\begin{equation}\label{eq:meth_prob}
\max_{\alpha \in \mathcal{U}} J(\alpha)\equiv \int_0^{\infty} \hat{r}(\hat{t}) \, \dfrac{\hat{p}(\hat{t})}{K+\hat{p}(\hat{t})}\, d\hat{t}.
\end{equation}
Since $J(\alpha)$ diverges, we actually consider overtaking optimality:
A solution is overtaking optimal if its performance index catches up with the performance index of any other solution (\cite{carlson_infinite_1991}, see \ref{S3_Text} for details).

Necessary conditions on optimal trajectories can be obtained by the Infinite Horizon Maximum Principle \cite{carlson_infinite_1991}, an extension of the well-known Pontryagin Maximum Principle.
Analysis of the Hamiltonian of the system of Eqs~\ref{eq:pdefnondim}-\ref{eq:rdefnondim} and the associated adjoint system shows that the optimal trajectory is a concatenation of bang arcs ($\alpha(\cdot)=0$ or $\alpha(\cdot)=1$) and possibly a singular arc corresponding to the optimal steady state $(\hat{p}(t),\hat{r}(t))=(\hat{p}_{opt}^*,\hat{r}_{opt}^*)$, that is, the steady state leading to the optimal growth rate $\hat{\mu}_{opt}^*$ in the new environment after the upshift (\ref{S3_Text}).
Moreover, from the Kelley condition~\cite{borisov_fullers_2000}, we can show that if the optimal trajectory has a singular arc, then it must enter this singular arc \textit{via} a chattering arc, \textit{i.e.}, with an infinite number of switches of $\alpha (\cdot)$ between 0 and 1 (\ref{S3_Text}).
The chattering arc is characterized by a switching curve $\hat{r}=\varphi(\hat{p})$ in the $(\hat{p},\hat{r})$-plane, which passes through $(\hat{p}_{opt}^*,\hat{r}_{opt}^*)$.
The switching curve divides the phase plane into two regions, such that $\alpha(t)$ switches to 0 when the system is in the region above $\varphi$ and to 1 when the system is below $\varphi$ (\ref{S3_Text} and Fig.~\ref{fig:optimalcontrol}).

The above results have led to the conjectured optimal solution of Eq.~\ref{eq:optcontrol}.
In parallel, we numerically solved the problem of Eq.~\ref{eq:meth_prob} by a direct method using the \texttt{bocop} software~\cite{bonnans_bocop_2012}.
A time discretization allows the optimal control problem to be transformed into a nonlinear optimization problem,
solved here by interior point techniques.
A discretization by a Lobatto IIIC formula (6th order) was used with 4000 time steps, and the relative tolerance for the NLP solver was set to $10^{-14}$.
The optimal trajectories thus obtained are composed of a chattering arc followed by a steady state corresponding to the singular arc (Fig.~\ref{fig:optimalcontrol}).
The switching curve $\varphi(\hat{p})$ was computed from numerical simulations with different initial conditions.

\subsection*{Specification and analysis of control strategies}

As described in the \textit{Results} section, we are interested in control strategies satisfying the following conditions:
\begin{description}
\item[(C1)] The control laws are static functions of the system variables (as opposed to, for instance, functions that depend on derivatives or integrals of the variables).
\item[(C2)] For any given constant environment $E_M$, they drive the self-replicator system towards a unique stable steady state that is not trivial, \textit{i.e.}, with nonzero growth rate.
\item[(C3)] This steady state corresponds to the optimal steady state $(\hat{p}_{opt}^*, \hat{r}_{opt}^*)$, allowing growth at the maximal rate $\mu^*_{opt}$.
\end{description}
It can be directly verified from the functions $f$, $g$, and $h$ defining the nutrient-only, precursor-only, and on-off control strategies (Eqs~\ref{eq:f}, \ref{eq:precursorstrategy}, and \ref{eq:stratswitch}) that they are indeed static functions of the system variables (or the system input, in the case of the nutrient-only strategy). 
Here we show that the other two conditions are also satisfied for all three strategies.

Following Eq.~\ref{eq:f}, the nutrient-only strategy is defined by $\alpha = f(E_M)$, so that $\alpha$ is constant after the upshift.
As shown above and in \ref{S1_Text}, this means that the system controlled by the nutrient-only strategy has a single nontrivial stable steady state (Condition~C2). 
In addition, in this case the optimal steady state is attained for $\alpha^*_{opt}$ defined as in Eq.~\ref{eq:alpha_mu_optimal}, and the following function $f$ therefore guarantees Condition~C3:
\begin{equation}
\label{eq:meth_f}
f(E_M) = \frac{E_M + \sqrt{K E_M}}{E_M + 2\sqrt{K\, E_M} + 1}.
\end{equation}
In \ref{S1_Text}, it is shown that Eq.~\ref{eq:meth_f} is the only definition of $f$ satisfying all conditions.
\ref{S1_Fig} shows a plot of $f(E_M)$ together with a biologically plausible Michaelis-Menten approximation (Eq.~\ref{eq:MMApprox}).

The full specification of the precursor-only strategy demands an expression for the function $g$ in Eq.~\ref{eq:precursorstrategy}.
Recall that Eq.~\ref{eq:alpha_mu_optimal_p} defines $\alpha_{opt}^*$ in terms of the precursor concentration $\hat{p}_{opt}^*$, which leads us to propose the following function $g$:
\begin{equation}
\label{eq:meth_g}
g(\hat{p}) = \frac{\hat{p}}{\hat{p} + \frac{K}{K + \hat{p}}(1+\hat{p})}.
\end{equation}
As shown in \ref{S1_Text} by computing the Jacobian, the system given by Eqs~\ref{eq:pdefnondim}-\ref{eq:rdefnondim} and \ref{eq:meth_g} has a single nontrivial stable steady state for any environment $E_M$ (Condition~C2).
Moreover, Eq.~\ref{eq:meth_g} guarantees this steady state to be optimal (Condition~C3).
This can be seen by noting that at steady state, $d\hat{r}/dt=0$ implies $\hat{r}^*=g(\hat{p}^*)$ (Eq.~\ref{eq:rdefnondim}).
In order for the self-replicator to attain a maximal growth rate at steady rate, Eq.~\ref{eq:alpha_mu_optimal_p} needs to be satified, which is the case for the above choice of the function $g$.
Like for $f$, Eq.~\ref{eq:meth_g} is the only choice for $g$ satisfying C1-C3.
\ref{S1_Fig} shows a plot of $g(\hat{p})$ together with a biologically plausible Hill approximation (Eq.~\ref{eq:HillApprox}).

The on-off control strategy is defined in Eq.~\ref{eq:stratswitch} and repeated below:
\begin{equation}
h(\hat{p}, \hat{r}) = 
\begin{cases}
0, \ \textrm{if} \ \hat{r} > g(\hat{p}),\\
1, \ \textrm{if} \ \hat{r} < g(\hat{p}), \\
\alpha_{opt}^*, \ \textrm{if} \ (\hat{p},\hat{r})=(\hat{p}_{opt}^*,\hat{r}_{opt}^*).
\end{cases}
\end{equation}
This strategy drives the system to a single steady state, because the $\hat{p}$-nullcline crosses the function $g(\hat{p})$ only once, as shown graphically in Fig.~\ref{fig:onoffresults}\textit{A}.
In \ref{S1_Text} we argue that this steady state is stable, by taking into account so-called sliding modes on the switching curve~\cite{filippov_differential_1988} (Condition~C2).
Moreover, the steady state coincides with the optimal steady state $(\hat{p}_{opt}^*,\hat{r}_{opt}^*)$ by construction, so that Condition~C3 is satisfied as well.
Fig~\ref{fig:ppGppsurface}\textit{A} shows a plot of $h(\hat{p}, \hat{r})$.

Note that since $h(\cdot)$ is discontinuous, numerical instabilities occur during simulations.
We therefore used the following continuous approximation of this function:
\begin{equation}
\frac{g(\hat{p})^{100}}{g(\hat{p})^{100} + \hat{r}^{100}}, \ \textrm{if} \ \hat{r} \neq g(\hat{p}).
\end{equation}
The approximation causes $\alpha$ to take intermediate values (instead of 0 or 1) just before reaching the optimal steady state in Fig~\ref{fig:onoffresults}\textit{C}.
For numerical simulations of the ODE system, we used the CVODE solver~\cite{cohen_cvode_1996} from SUNDIALS 2.6.2~\cite{hindmarsh_sundials_2005}.

\section*{Supporting Information}

%%%%%% FOR MANUAL REF%%%%%
\makeatletter
\newcommand{\manuallabel}[2]{\def\@currentlabel{#2}\label{#1}}
\makeatother
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\subsection*{S1 Text -- Model derivation and analysis}
\manuallabel{S1_Text}{S1~Text}

\subsection*{S2 Text -- Model parameters}
\manuallabel{S2_Text}{S2~Text}

\subsection*{S3 Text -- Solution of optimal control problem}
\manuallabel{S3_Text}{S3~Text}

\subsection*{S4 Text -- Kinetic model of the ppGpp system in \textit{Escherichia coli}}
\manuallabel{S4_Text}{S4~Text}

\subsection*{S1 Table -- Parameter values of self-replicator model}
\manuallabel{S1_Table}{S1~Table}

\subsection*{S1 Figure -- Simple control strategies for the self-replicator of bacterial growth}
\manuallabel{S1_Fig}{S1~Figure}

\section*{Acknowledgments}
The authors thank Eugenio Cinquemani for discussions and comments on the manuscript.

\nolinenumbers

%\section*{References}
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here.
%