### 2016-06-01 Chapters/authors_affil.tex

% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Nils Giordano \textsuperscript{1,3},
Francis Mairet \textsuperscript{2},
Jean-Luc Gouzé \textsuperscript{2},
Johannes Geiselmann \textsuperscript{1,3,*},
Hidde de Jong \textsuperscript{3,*}
\\
\bigskip
\bf{1} Université Grenoble Alpes, Laboratoire Interdisciplinaire de Physique (CNRS UMR 5588), 140 rue de la Physique BP 87, 38402 Saint Martin d'Hères, France
\\
\bf{2} Inria, Sophia-Antipolis Méditerranée research centre, 2004 route des Lucioles, BP 93, 06902 Sophia-Antipolis Cedex, France
\\
\bf{3} Inria, Grenoble - Rhône-Alpes research centre, 655 avenue de l'Europe, Montbonnot, 38334 Saint Ismier Cedex, France
\\
\bigskip

* Corresponding authors with equal contributions:\\
Hidde.de-Jong@inria.fr, Hans.Geiselmann@ujf-grenoble.fr