### 2016-06-01 Chapters/preambule.tex

%\usepackage{epic}
\usepackage[english]{babel}
%\usepackage{appendix}
%\usepackage{lipsum}
%\usepackage{natbib}

\usepackage[hypcap]{caption}

%%%%% Math Stuff
\usepackage{amsmath}
\usepackage{amsfonts}
% New operator
\newcommand{\argmax}{\operatornamewithlimits{argmax}}
% Theorems
\usepackage{amsthm}
\newtheorem{prop}{Proposition}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% Réglage des interlignes
%\usepackage{setspace}
%\onehalfspacing

% For roman numbers
%\renewcommand\thesection{\Roman{section}}

% New colors
%\usepackage{color} % for color stuff…
%\definecolor{lightred}{rgb}{1, 0.3, 0.3}

\usepackage{hyperref}
\hypersetup{
unicode=false, % non-Latin characters in Acrobat’s bookmarks
pdftoolbar=true, % show Acrobat’s toolbar?
pdfmenubar=true, % show Acrobat’s menu?
pdffitwindow=false, % window fit to page when opened
pdfstartview={FitH}, % fits the width of the page to the window
pdftitle={Environment dynamic structures optimality in cellular economics}, % title
pdfauthor={Nils Giordano}, % author
pdfsubject={Machinery optimal control article}, % subject of the document
pdfcreator={pdflatex}, % creator of the document
pdfproducer={LateX}, % producer of the document
pdfkeywords={gene expression machinery} {optimal control} {cell composition}, % list of keywords
pdfnewwindow=true, % links in new window
colorlinks=true, % false: boxed links; true: colored links
linkcolor=blue, % color of internal links (change box color with linkbordercolor)
citecolor=blue, % color of links to bibliography
filecolor=magenta, % color of file links
urlcolor=cyan % color of external links
}