### 2016-06-01 theory.tex

\chapter{Dynamical Allocation of Cellular Resources as an Optimal Control Problem}

\section*{Summary of chapter \thechapter}
In this section, we present the topic of microorganism growth, in particular the recent focus which is made on resource allocation, and discuss some of the limitations of those studies.
We describe the self-replicator approach, and how we're going to apply it in dynamical conditions.

\selectlanguage{french}
\section*{Résumé du chapître \thechapter : [titre français]}
dans cette section blablabla
\selectlanguage{english}