### 2016-07-25 acknowledgements.tex
Remove useless pages, reorganize some parts

\cleardoublepage
\begin{acknowledgements}
(remerciements ici)
\end{acknowledgements}