### 2016-12-22 Chapters/introduction.tex
Introduction 1.2 and first part of 1.3

\chapter{Introduction}

\textit{There's an infinity of things that have never been done before, and most of those things are not worth doing.} -- Jeremy Fox, \textit{Dynamic Ecology (2016)}~\cite{fox_how_2016}

\selectlanguage{french}
\section*{Résumé en français}

Dans cette section... (env. 1-2 pages)
\selectlanguage{english}

\begin{center}
\noindent\rule{4cm}{0.1pt}
\end{center}

\section{Context}
\label{sec:context}

\subsection{Self-replication is a resource allocation problem}

At every moment of our lives, we are surrounded by billions of microscopic organisms.
They sustain themselves by collecting energy from their environment, in the form of organic matter, highly reactive chemicals, or light~\cite{madigan_biology_2006,schaechter_microbe_2006}.
While most of our cells are kept in a stable and friendly internal environment (temperature, oxygenation, nutrient abundance, ...), microorganisms have evolved to constantly adapt their physiology to strongly variable environments.
Not without a certain success, because microbes have been found to survive almost everywhere, sometimes in the strangest places that were initially thought unsuitable for life~\cite{rothschild_life_2001,nicholson_transcriptomic_2012,madigan_biology_2006,schaechter_microbe_2006}.
Such ubiquity translates a strong potential for adaptation.
Even the model organism \textit{Escherichia coli}, a bacteria first isolated from our intestinal track, is capable of growing on dozens of different carbon sources through changes in the expression of thousands of genes\cite{zimmer_microcosm:_2009}.

Microorganisms, along with all living cells, are mainly constituted of polymers (DNA, RNA, Proteins) made of many repeated sub-units (nucleotides, amino acids).
The information about their sequences is stored in the form of genes on the DNA.
Gene expression is the process by which this information is used in the synthesis of a functional final product (protein or RNA).
The gene expression machinery (itself constituted of RNA and proteins) synthesizes the gene products after binding to gene promoters, which are target sequences on the DNA.
In a given cell, they are roughly as many different target sequences as genes, which ensures that not all products are synthesized at the same rate.
Furthermore, some of these products specifically bind some promoters and stimulates or represses the promoter activity, hence, the gene expression.
They are called transcription factors and allow, among other mechanisms, the regulation of cell composition.

The reorganization of gene expression controls the abundance of enzymes and ribozymes that catalyzes the biochemical reactions allowing the cell to perform different functions.
For instance, assimilating a given nutrient involves the synthesis of the corresponding uptake protein and of the metabolic enzymes converting the nutrients into reductive power, nucleotides and amino acids~\cite{schaechter_microbe_2006}.
These precursors are then used to produce new macromolecules, closing the loop of self-replication.
In order to optimize the overall process in different environmental conditions, the cell must adjust the abundance of every enzyme involved in this network of biochemical reactions.
By doing so, the cell actually resolves a resource allocation problem, where a common wealth (precursors) has to be distributed over thousands of potential beneficiaries (proteins and other macromolecules).

\subsection{Optimizing biomass production yields a competitive advantage}

Which resource allocations are optimal for a living cell?
Systems capable of Darwinian evolution sustain themselves on the long term by \textit{fitting} to their environment.
Living cells with the best \textit{fitness} are more likely to survive, replicate, and through long-term competition, eliminate their siblings.
The optimization criteria that quantitatively measure \textit{fitness} are called \textit{fitness factors}.
Whether they apply at the species, the population, the organism, or even the gene level has long been debated~\cite{dawkins_selfish_1976}.
An optimal resource allocation would be a gene expression scheme that maximizes the fitness factors.
But since fitness factors have to take into account the ability to survive, reproduce, and compete with other organisms in a given (sometimes changing) environment, they are often context-dependent and highly difficult to identify.

For this reason, there is no consensus about what constitute the main fitness factors of microorganisms.
Competition for common resources seems to favor the maximization of offspring, in the sense that a replicator with more descendants will outnumber its competitors in the long term~\cite{dawkins_selfish_1976}.
It is unclear, however, how this general principle translates into their physiological behaviors.
For instance, the use of genome-scale metabolic network models have shown that the metabolism maximizes the production either of biomass or available energy (ATP) depending on the environmental condition~\cite{schuetz_systematic_2007}.
But we can also found different strategies when the biomass production is at play.
Bacteria can either optimize their production yield (taking its time to make the most of the nutrient) or their production rate (growing quickly while wasting part of the nutrient potential), the later being favored in fluctuating or poor environments~\cite{frank_tradeoff_2010,maclean_tragedy_2008}, a result that has also been confirmed in yeast~\cite{schuster_maximization_2008}.
Overall, it seems that cell optimality drifts in a multidimensional space and involves making several trade-offs that strongly depend on the environmental context~\cite{schuetz_multidimensional_2012}.

Nevertheless, numerous studies have shown that considering growth rate maximization confers a good predictive power of microorganism physiology.
When cultivated in minimal medium with acetate or succinate, \textit{Escherichia coli} uptake rates are correctly predicted by assuming that the metabolic network operates in a mode that maximizes growth rate~\cite{edwards_silico_2001}.
This result is not general though, and depends on the carbon source and strain used~\cite{ibarra_escherichia_2002}.
But even in those cases, the metabolic networks evolve rapidly towards growth-rate maximization if a constant nutrient-providing environment is maintained~\cite{ibarra_escherichia_2002}.
The mutants obtained present modications in the regulation of the synthesis of metabolic enzymes~\cite{lewis_omic_2010}, indicating that not only the efficiency of the enzymes is affected, but also their abundance and thus the cellular composition.
Overall, with the currently available knowledge and its limitations (discussed in section~\ref{chap:discussion}), growth-rate maximization sounds like a rational choice to understand the growth strategies of microorganisms~\cite{molenaar_shifts_2009}.

\subsection{Growth laws are universal strategies employed by microorganisms}

One of the first attempt to formalize microorganism physiology into elegant, fundamental laws was the discovery of Monod's law more than sixty years ago~\cite{monod_growth_1949}.
This empirical relationship states that the growth rate $\mu$ [\texttt{div.h\textsuperscript{-1}}] of \textit{Escherichia coli} is a hyperbolic function of the concentration of the limiting nutrient $C$ [\texttt{mol.L\textsuperscript{-1}}], such that
\begin{equation}
\label{eq:monod_law}
\mu = \mu_K \frac{C}{K_C + C}
\end{equation}
with $\mu_K$ and $K_C$ two constants depending on the quality of the nutrient (Fig~\ref{fig:monod_law}).
It is remarkable that the growth rate, a parameter that depends on the use and production of thousands of proteins, can be so easily predicted by a simple relation.
What is even more remarkable is that this relation is rather universal.
Despite some adjustments (the Droop model being a well-known example~\cite{droop_thoughts_1973}), the relation holds for most microorganisms: the growth rate increases with the abundance of the limiting nutrient, up to a point beyond which they cannot grow any faster~\cite{koch_why_1988}.

\begin{figure}[tb]
\centering
\includegraphics[height=5cm]{./Fig/Chapter1/monod_law.eps}
\caption{
\textbf{Monod law, reproduced from~\cite{monod_growth_1949}.}
Monod has been a pioneer in the formalization of microbial growth into fundamental, coarse-grained relationships.
This figure displays the growth rate (in number of divisions per hour) of \textit{Escherichia coli} in a synthetic minimal medium at 37$^{\circ}$C containing different concentrations of glucose (a common carbon source).
As the concentration of glucose increases, the growth rate of \textit{E. coli} does so by following the hyperbolic relationship described in Eq~\ref{eq:monod_law}.
Solid line is drawn with $\mu_K = 1.35$ div.h\textsuperscript{-1}, and $K_C = 0.22 \cdot 10^{-4}$ mol.L\textsuperscript{-1}.
}
\label{fig:monod_law}
\end{figure}

The production of every component of the cell is controlled by a complex network of regulatory interactions.
But an obvious constraint is that at steady state, the synthesis rate of all individual components must be proportional to the growth rate in order to compensate for growth dilution~\cite{monod_growth_1949}.
Many physiological parameters (like mass of DNA, RNA and proteins) are thus functions of the growth rate alone, regardless of the environmental conditions~\cite{schaechter_dependency_1958,bremer_modulation_1996}.
These so-called growth laws were carefully measured~\cite{bremer_modulation_1996} and are still used today in the quantitative understanding of growth control in microorganisms~\cite{ehrenberg_mediumdependent_2012}.
Recently, the topic was revitalized through the work of Scott \textit{et al}~\cite{scott_bacterial_2011}.
They focused on the ribosome concentration, a physiological parameter that was long known to vary linearly with the growth rate in microorganisms (Fig~\ref{fig:scott_rnaprot}).
By measuring it under different environmental perturbations of the protein synthesis machinery, they built a coarse-grained model describing proteome resource allocation~\cite{scott_emergence_2014}.
It has allowed to show that, whatever the details of the molecular implementation (differing between organisms), this growth law emerges from the underlying principles of robustness and optimization imposed by natural selection, especially growth-rate maximization in every conditions~\cite{scott_emergence_2014}.

\begin{figure}[tb]
\centering
\includegraphics[height=7cm]{./Fig/Chapter1/scott_rnaprot.eps}
\caption{
\textbf{The growth law of ribosomal abundance (figure reproduced from Fig~S1 in~\cite{scott_interdependence_2010}).}
For a variety of carbon nutrients and their corresponding steady-state growth rate, the RNA/protein ratio is linearly correlated with the growth rate, a relation that holds for many species of microorganisms.
This ratio is correlated with the fraction of ribosome-affiliated proteins, and therefore with the relative abundance of ribosomes in the cell.
}
\label{fig:scott_rnaprot}
\end{figure}

\subsection{Static versus dynamical perspective on growth}

The growth laws cited above apply at steady state, where all intensive properties of the cell are time-invariant~\cite{schaechter_microbe_2006,fishov_microbial_1995}.
It means that the properties independent of the cell volume or mass (temperature, concentrations, ...) are constant over time, even though the cell is still growing.
It requires that the components of the cell "increases by the same factor over a time interval", which have motivated the creation of the term balanced growth~\cite{campbell_synchronization_1957}.
Experimentally, this growth scenario has been used as a standard because it improves reproducibility, in that samplings are no longer time-sensitive~\cite{schaechter_microbe_2006}.
It can be easily achieved in the laboratory either in continuous culture, where the substrate is continually supplied, or in batch conditions if the substrate is in high excess (Fig~\ref{fig:growth_curve}).
This approach has been beneficial to mathematical modeling because considering the system at steady state reduces the complexity of the underlying dynamical systems, allowing to build and analyze huge genome-scale models that encapsulate the enzymatic diversity of living cells~\cite{orth_what_2010}.

\begin{figure}[p]
\centering
\includegraphics[height=6cm]{./Fig/Chapter1/growth_curve.eps}
\caption{
\textbf{The different phases of a typical growth curve}.
During a typical batch growth scenario, biomass accumulates (black thick line) whereas nutrients are consumed until depletion (green dashed line)~\cite{schaechter_microbe_2006}.
(A) The lag phase is a variable period of time during which the organism adapt to the new medium~\cite{swinnen_predictive_2004}.
It is hard to study experimentaly and is known to be affected by the pre-culturing history of the strain~\cite{ng_damage_1962,dufrenne_effect_1997,shaw_effect_1967}, the magnitude and the rate of the change between the past and present environments~\cite{mcmeekin_predictive_2002}, and other hard-to-control environmental conditions~\cite{cheroutre-vialette_application_2002}.
(B) The steady-state (or balanced-growth) phase is characterized by an exponential production of biomass.
Its characteristics are time-invariant and reproducible, which has made it a standard for microbial growth studies~\cite{schaechter_microbe_2006}.
This phase can be extended for hundreds of generations in continuous cultures, by the constant addition of fresh medium~\cite{wang_robust_2010}.
As represented here, nutrients are quickly depleted in batch conditions, which does not allow to maintain steady-state growth for a long period of time.
(C) The stationary phase occurs after the depletion of the limiting nutrients in the medium~\cite{chubukov_environmental_2014,schaechter_microbe_2006}.
In natural conditions, microorganisms usually encounter poor media and spend most of their time in stationary phase~\cite{mcarthur_microbial_2006,menge_nitrogen_2012,hobbie_microbes_2013}.
Some species which have evolved long-term resistance mechanisms, like sporulation~\cite{stragier_molecular_1996} or cannibalism~\cite{gonzalez-pastor_cannibalism:_2011}, can survive particularly long stationary phases.
}
\label{fig:growth_curve}
\end{figure}

Although balanced growth is convenient from an experimental and theoretical point of view, it is widely admitted that microorganisms rarely encounter this state in nature~\cite{schaechter_microbe_2006}.
Steady-state growth requires stable conditions over a long period of time, but microorganisms live in
environments where the key elements, such as carbon, nitrogen or phosphorus, are quickly depleted by the competitors as soon as they become available~\cite{mcarthur_microbial_2006,menge_nitrogen_2012,hobbie_microbes_2013}, a consideration that also holds for the lab strains on which growth laws are studied~\cite{savageau_escherichia_1983,savageau_demand_1998,blount_unexhausted_2015,vanelsas_survival_2011}.
Why would microorganisms be optimal for a state they have barely encountered during their evolution?
Are we missing something by studying in stable, unchanging conditions, systems that were in fact selected to cope with environmental variations?

\section{Problem statement}

Microbial growth is essentially a resource allocation problem that can be summarized in simple, fundamental growth laws.
Though established empirically, those laws are elegantly explained if we consider that microorganisms behave in way that optimizes their biomass production.
However, these laws describe growth in stable environments, but it is hard to imagine how natural selection could have made microorganisms optimal for conditions they rarely encounter.
This motivates the study of growth laws through a dynamical perspective, that is, in a changing environment, and leads us to the following problem statement: \textbf{Which strategies microorganisms follow to dynamically reallocate their resources after a change in the environment?}
The development of this question prompts the following investigations.

What are the best strategies of resource allocation if, like for the steady-state growth laws, the biomass production is kept as the only optimization criteria?
Through answering this question, we could assess to which extend the features of dynamical optimization differ from the steady-state optimization.
In other words, it would test our verbal hypothesis that motivated a dynamical perspective on growth: the fact that we expect evolutionary pressure to apply differently between dynamic and stable conditions.
This requires to translate the problem in the mathematical world, through the development of a proof-of-concept model of resource allocation in bacterial cells, and the theoretical determination of the optimal dynamical allocation of resources following a change in the environment.

How could one measure empirically the resource allocation during a growth transition?
The model will help us to establish actual resource allocation strategies that optimize biomass production during an environmental change.
However, it will not be able to tell us if microorganism actually care to optimize such a criterion.
We need though the possibility to measure biomass accumulation and changes in resource allocation when a shift is made in the environment.
This would require a way to dynamically track the abundance of key macromolecules in living cells, while establishing an experimental set-up that temper the inherent variability of dynamical experimental studies of microbial growth.

\section{Related work}

\subsection{Modeling growth of microorganisms}

My recent experience as a teacher taught me that, as of today, a lot of students in biology do not like mathematics very much.
Since mathematics is the main language of modeling, this has often made it difficult for me to convince them that modeling is of key importance for biologists.
However, while mathematics is the language, it is not the essence of modeling.
Generally speaking, a model depicts and simplifies reality.
Maps, sketches, or pictures satisfy this definition, as do graphs, sets of equations, or even the DNA sequences stored in a text file on a computer.
In other words, depicting and simplifying reality makes you a modeler, not so much drawing equations on a board.
But this does not change the fact that mathematics are powerful tools for formulating and analyzing models~\cite{servedio_not_2014,mcgill_calm_2013}.

What makes mathematical modeling so useful in microbial growth studies?
Evolutionary biology is a good instance of field that has been resorting for a long time on mathematical modeling~\cite{servedio_not_2014}.
The large time and population scales at which evolutionary processes occur truly impede most of the experimental work, despite some attempts on organisms with a short generation time~\cite{elena_evolution_2003}.
Besides, microbiology is a field particularly prone to experimental study.
A growth curve as presented in Fig.~\ref{fig:growth_curve} can be made in a day, so the limitations of evolutionary biology should not apply to microbial growth studies.
Why care to mimic reality on a computer when we can just directly bend it to our will~\cite{hillis_why_1993}?
The last decades have strongly challenged this viewpoint: as knowledge and data on microbes accumulated, the need for a systemic understanding of biology was more and more pressing~\cite{alon_introduction_2006,kremling_systems_2013}.

Microbial physiology results from the interplay of thousands of chemical reactions that are not necessarily relevant in any given situation~\cite{schaechter_microbe_2006}.
These reactions occurs on short time scales, and are controlled by several layers of regulatory mechanisms that can only be understood through evolutionary considerations.
As a consequence, even the simplest verbal hypothesis can resist direct empirical testing.
But abstracting away complexity is the purpose of mathematical models.
The overwhelming number of variables and time scales can be dealt with by choosing the correct framework.
When working on a model, the microbe behaviors are abstracted into a world of clearly stated rules, where logical reasoning can be applied~\cite{servedio_not_2014,mcgill_calm_2013}.
Predictions can be made, "unpacked" into our real world, and confronted with experimental testing.

Modeling was particularly helpful to unveil how metabolic networks operate.
For an increasing number of microorganisms, we are now able to draw quasi-exhaustive maps of their metabolic reactions.
For instance, the most recent genome-scale reconstruction of \textit{E. coli} metabolism contains 1366 genes, 2251 metabolic reactions, and 1136 unique metabolites~\cite{orth_comprehensive_2011}.
Such a number of variables might seem overwhelming, and making sense of these knowledge is rarely straightforward.
Which reactions are important, and in which environmental conditions?
Can we predict how perturbations will affect a given metabolic network?
Do fundamental regularities exist between different metabolic networks, from different species?

This apparent complexity does not impede a mathematical studies of metabolic networks.
Constraint-based modeling is a framework that abstracts away the unknown kinetics and represents the metabolic reactions by steady-state fluxes to which physico-chemical constraints can be applied, e.g. compartmentalization, mass conservation, molecular crowding, and thermodynamic directionality~\cite{ebrahim_cobrapy_2013}.
In other words, through constraints imposed to the metabolic networks, it provides a space of admissible flux distributions, eliminating reactions that are unlikely to occur in the environmental condition of interest.
This space is often further reduced by selecting the flux distributions optimizing a specified objective function (e.g., the growth rate of the cell).
Such an approach, initially developed through flux balance analysis (FBA)~\cite{orth_comprehensive_2011,palsson_systems_2011}, has been shown to correctly predict many behaviors of the metabolic network of \textit{E. coli}~\cite{varma_stoichiometric_1994,edwards_silico_2001}.
It also proved capable of predicting its long-term recovery through genetic evolution after a gene deletion~\cite{fong_metabolic_2004}.

More direct modeling approaches, in the form of large kinetic models, have also been helpful to understand growth-related processes.
Through the construction of a model of 47 differential equations and 193 parameters, Kotte \textit{et al} showed how metabolic fluxes are sensed at different locations of the central carbon metabolism, and integrated in a global cell response by the coupling of enzymatic and transcriptional regulations~\cite{kotte_bacterial_2010}.
Such models are often sufficiently detailed to be used as \textit{in silico} testbed to investigate and uncover specific molecular mechanisms~\cite{peskov_kinetic_2012}.
The most emblematic instance of that kind of approach is probably the recent whole-cell model of \textit{Mycoplasma genitalium}~\cite{karr_whole-cell_2012}.
By aggregating all the available knowledge from more than 900 publications, Karr \textit{et al} constructed a dynamical model accounting for all the annotated gene functions of this pathogenic organism, and successfully used it to investigate unobserved molecular mechanisms and guide novel experimental analysis.
While such approach is not yet applicable to all kind of organisms, it genuinely demonstrates how the increase in computational power could one day make at hand \textit{in silico} experimentation.

But while it represents their strength, the granularity of the models cited above is also their main weakness.
Despite this apparently rich available knowledge, precise measurements of \textit{in-vivo} kinetic parameters still represents a technological bottleneck~\cite{park_metabolite_2016,bennett_absolute_2009,buscher_cross-platform_2009}.
Constraint-based modeling overcomes this limitation by using a mathematical approach that does not require any knowledge about the reaction kinetics.
The drawback is that they highly depend on the constraints used to reduce the space of admissible solutions, e.g. the choice of the objective function that could sometimes be tricky~\cite{kauffman_advances_2003}.
Kinetic models however are particularly sensible to missing parameters.
While the so-called free parameters are often collectively fitted to available data~\cite{jaqaman_linking_2006,mendes_non-linear_1998}, this is known to produce large parameter uncertainty~\cite{cho_experimental_2003,brodersen_characterization_1987,rodriguez-fernandez_hybrid_2006}, which can significantly affect the predictions of the model~\cite{gutenkunst_universally_2007,ingram_network_2006,mayo_plasticity_2006}.

Such limitations often motivate the construction of simpler, coarse-grained models.
The approach is dramatically different: instead of aggregating the available knowledge into a highly detailed model, the modeler is concerned about carefully filtering this information to keep the model as simple as possible.
They generally consider a high level of abstraction of the cell functions, and are particularly valuable when looking for universal laws or principles~\cite{scott_bacterial_2011,scott_interdependence_2010,scott_emergence_2014}.
They can take the form of minimal core models that focus on a given aspect of growth, while still abstracting away the molecular details (see \textit{e.g.}, ~\cite{spiesser_size_2012}).
Such models can also be used as proof-of-concept to submit verbal hypothesis to the logic and rigor of mathematical reasoning~\cite{servedio_not_2014}.
For instance, a simple proof-of-concept model was used to uncover principles that lead to overflow metabolism, a mechanism by which microorganisms switch to inefficient metabolic pathways when growing at high nutrient availability~\cite{molenaar_shifts_2009}.
This paradoxal and widely spread mechanism~\cite{dijken_kinetics_1993,vemuri_overflow_2006,mckeehan_glycolysis_1982,hsu_cancer_2008} was shown to be easily explained by the trade-off between the efficiency of a pathway and the cost of synthesis of its enzymes.
Overall, these models have several advantages for the purpose of our study: they clearly state and do not lose track of their assumptions, and are sufficiently tractable to allow access to complex mathematical tools (as will be required for dynamical optimization).

%It is also the preferred approach when one wants to broaden its access to mathematical tools~\cite{vandenberg_optimal_1998}, most of them being unhelpful with the high dimensionality of biological models.
%As we will describe in section~\ref{sec:approach}, this represents the modeling strategy that was used in this thesis.

%\begin{figure}[tb]
%\centering
%\includegraphics[width=\textwidth]{./Fig/Chapter1/molenaar_overflow.eps}
%\caption{
%\textbf{The strategy of overflow metabolism, reproduced from~\cite{molenaar_shifts_2009}.}
%When increasing the concentration of substrate, the metabolism of microorganisms switches toward the use of inefficient pathways.
%This paradoxical behavior does in fact maximize the growth rate if we take into account the costs and benefits of enzyme synthesis, in particular the fact that pathways for inefficient catabolism are shorter and therefore cheaper to produce.
%(A) Simple view of microbial metabolism, with the two competing efficient and inefficient pathways.
%(B) The switching from efficient to inefficient catabolism when the concentration of nutrients, and hence the growth rate, increases.
%}
%\label{fig:molenaar_overflow}
%\end{figure}

\subsection{Measuring growth of microorganisms}

\textbf{Either to build models or test their predictions, data on microbial growth have to be acquired.}
Measuring growth = measuring biomass accumulation, over time, and get relevant information on the cell inner state.
We have a lot of data on this.

\textbf{Measuring population-wide parameters is convenient and provides robust data.}
Population measurement is easy to implement, and some things can only be measured at the population level (so far).
But small unsynchronized effects can be hidden since we mostly work with the mean.
Despite most of the knowledge we have on microorganism growth are from population measurements, like the RNA/DNA/etc ratios, the growth laws, etc.
Extensive work of Bremmer and Dennis.
Big discoveries of the last decades that rely on population measurements ? (cite 2 or 3)

\textbf{Single-cell measurements give access to data and behaviors that are hidden at the population level.}
Single-cell measurements lay on technologies like microscopy, fluorescence, FACS...
More interesting since we can use fluorescence to not just look at the cell but also measure interesting parameters (like gene expression) at the single-cell level.
Information on the localisation of DNA, ribosomes [Bakshi].
Variability in physiology, persistors, antibiotics resistance, etc [citations]

\textbf{Batch growing conditions are convenient and close to the natural conditions of bacteria, but they raise several problems.}
Both approaches can be performed either in batch or continuous conditions.
Batch conditions is the most straight forward in lab: you put sugar in a flask, they grow until nutrient is depleted.
Implemented in flash or petri-microscopy.
Resist to contamination.
It is close to natural conditions, and industrial conditions.
Has even been used to study long-term evolution of E. coli.
But the system is poorly control, mostly because the history of the cells play a huge role in the waking process.

\textbf{Continous cultures allow to precisely control the growing conditions.}
Continuous culture consists into constantly providing nutrients to a culture, and maintaining it in a steady-state for a long time.
Implemented in bioreactor [citation] or microfluidic devices [mother machine citation].
Sensible to contaminations so hard to maintained (that's why industry ignore it so far).
Does not correspond to natural conditions but the system is fully controlled: perturbations can be applied and a lot of parameters can be monitored.
Possibility to let the cell long-enough in steady state to "forget" strain history.
Most suitable if one need to have full control on its system.

\textbf{In all these practical conditions, one can choose to either measure cells at steady-state or in dynamic.}
Finally, most of the work cited above are based on steady-state studies.
But there are also work in dynamic.
Levy showed that... in yeast.
Alon in Coli.
After decades working at steady-state, Bremmer and Dennis argued we should measure things during transitions.
de Jong / Gerosa showed that the gene expression machinery has its own dynamic that should be taken into account when measuring gene expression.
The evolutionary relevance of growth transitions has been emphasized in numerous others studies [citations lag phase, last generation before stationary, ...].

\section{Approach}
\label{sec:approach}

We first tackle the problem of creating a mathematical framework for studying environmental transitions.
\textit{Building upon the work cited above, we create a simple self-replicator model that allocate resources between two sections.}

The goal is to identify what would be an ideal transition.
\textit{We all have an intuitive idea of what should be a good transition: it should be fast, robust, etc... But it is unclear how this relates to the fitness factor that shaped the regulatory mechanisms of microorganisms.}

We describe dynamical optimality by using optimal control theory.
\textit{Brieve explanation about optimal control theory and how it is used in economic or engineering to optimize or understant dynamical processes.}

All of this is described in chapter 2.

Then, after we know what to look for, we try to actually measure how resource allocation occurs in a model organism: \textit{Escherichia coli}.
\textit{We justify the use of coli and introduce the gene expression fluorescent reporter.
We made a strain with fluorescent ribosomes, and studied our experimental framework in different transitions, at the population and the single cell level.
We justify the use of both by the fact that population is easy for "screening", and single cell allow for a better resolution and take care of the "desynchronization" biais that could occur.}

This, is described in chapter 3.