### 2017-01-02 acknowledgements.tex

\cleardoublepage
\selectlanguage{french}
\begin{acknowledgements}
(remerciements ici)
\end{acknowledgements}
\selectlanguage{english}